# Example Territory Maps {background-color="#A1D6F4"}

A review of the 3 largest New York Home Insurance Companies

## No. 3: Chubb (626) - Chubb National Insurance Company (10052) {background-color="#A1D6F4"}

Perils:

- Hurricane Wind
- Non-Hurricane Wind
- All Other Perils

Territories:

- A gnarly mix of county, town, ZIP codes, and islands

For example, New York County is split into three pieces:

- Governors Island - Territory 97
- Fort Charles - Territory 05
- Rest of County - Territory 18

I'm not even going to try...

## No 2: Allstate Insurance Group (8) - Allstate Vehicle and Property Insurance Company (37907) {background-color="#A1D6F4"}

ZIP Code-based territories

12 perils. Similar to the ROC with the additional of PC-11 (Hurricane)

::: {.notes}
Their esurance company in Nevada uses 0.05, 0.1, 0.2, 0.4 grid  SERFF ESUR-129867328 Page 920

:::

## Allstate - Weather Water {background-color="#A1D6F4"}

```{r Allstate NY Weather Water}
#| echo: true
#| code-fold: true
#| code-summary: "Show code"

# Read in Allstate factor table
allstate_ny <- read_csv("./references_private/02_Allstate_Factors.csv",
                        col_names = c("zip",
                                      "fire1", "fire2", "brush_fire", "water_weather",
                                      "water_nonweather", "wind", "hail", "lightning",
                                      "theft", "liability", "hurricane", "other"),
                        skip = 1)

zip_2024  <- read_sf("data_in/geo/tl_2024_us_zcta520/tl_2024_us_zcta520.shp")

# Join factors to ZIP shapefile
zip_allstate_ny <- zip_2024 %>%
  inner_join(allstate_ny, by = c("ZCTA5CE20" = "zip"))

# Set up a color palette
pal <- colorBin(
  palette = c("#4575b4", "#91bfdb","#e0f3f8","#fee090","#fc8d59","#d73027"),
  domain = zip_allstate_ny$water_weather,
  bins = c(0.6, 0.8, 1.0, 1.20, 1.4, 1.6, 1.8),
  na.color = "transparent"
)

# Plot on top of Open Street Maps basemap
leaflet(zip_allstate_ny) %>%
  addTiles() %>%
  addPolygons(
    fillColor = ~pal(water_weather),
    color = "white",
    weight = 1.0,
    fillOpacity = 0.7) %>%
  addLegend(pal = pal, values = ~water_weather, title = "Weather Water")
```

## Allstate - Wind {background-color="#A1D6F4"}

```{r Allstate NY Wind}
#| echo: true
#| code-fold: true
#| code-summary: "Show code"


# Set up a color palette
pal <- colorBin(
  palette = c("#4575b4", "#91bfdb","#e0f3f8","#fee090","#fc8d59","#d73027"),
  domain = zip_allstate_ny$wind,
  bins = c(0.6, 0.8, 1.0, 1.20, 1.4, 1.6, 1.8),
  na.color = "transparent"
)

# Plot on top of Open Street Maps basemap
leaflet(zip_allstate_ny) %>%
  addTiles() %>%
  addPolygons(
    fillColor = ~pal(wind),
    color = "white",
    weight = 1.0,
    fillOpacity = 0.7) %>%
  addLegend(pal = pal, values = ~wind, title = "Wind")
```

## Allstate - Theft {background-color="#A1D6F4"}

```{r Allstate NY Theft}
#| echo: true
#| code-fold: true
#| code-summary: "Show code"


# Set up a color palette
pal <- colorBin(
  palette = c("#4575b4", "#91bfdb","#e0f3f8","#fee090","#fc8d59","#d73027"),
  domain = zip_allstate_ny$theft,
  bins = c(0.5, 0.75, 1.0, 1.25, 1.5, 1.75, 2.00),
  na.color = "transparent"
)

# Plot on top of Open Street Maps basemap
leaflet(zip_allstate_ny) %>%
  addTiles() %>%
  addPolygons(
    fillColor = ~pal(theft),
    color = "white",
    weight = 1.0,
    fillOpacity = 0.7) %>%
  addLegend(pal = pal, values = ~theft, title = "Hurricane Factor")
```


## Allstate - Hurricane {background-color="#A1D6F4"}

```{r Allstate NY hurricane factors}
#| echo: true
#| code-fold: true
#| code-summary: "Show code"

# Set up a color palette
pal <- colorBin(
  palette = c("#4575b4", "#91bfdb","#e0f3f8","#ffffbf","#fee090","#fc8d59","#d73027"),
  domain = zip_allstate_ny$hurricane,
  bins = c(0, 0.5, 1.5, 2.5, 4.0, 5.0, 6.0, 7.5, 10.0),
  na.color = "transparent"
)

# Plot on top of Open Street Maps basemap
leaflet(zip_allstate_ny) %>%
  addTiles() %>%
  addPolygons(
    fillColor = ~pal(hurricane),
    color = "white",
    weight = 1.0,
    fillOpacity = 0.7) %>%
  addLegend(pal = pal, values = ~hurricane, title = "Hurricane Factor")
```


## State Farm Group (176) {background-color="#A1D6F4"}

State Farm Group (176)

State Farm Fire And Casualty Company (25143)

Ranked #1 in NY ($1.2B in 2024 Direct Written Premium)

- Granularity: "Location Rating" Grid
  - Decimal degrees to two decimal places
- 2 Perils:
    - Hurricane 
    - Non-Hurricane
- Source: 4,637 PDF pages of data

::: {.notes}
Extracted using Tabula
Excel get data from PDF also works well
:::

## State Farm Fire And Casualty Company (25143) - Non-Hurricane Factors {background-color="#A1D6F4"}

```{r New York State}
#| echo: true
#| code-fold: true
#| code-summary: "Show code"

# Starts extracting the table from PDF
sf_ny <- read_csv("./references_private/StateFarmNY_20200715.csv", col_names = c("GRID_ID", "HURR_FCTR", "NONHURR_FCTR"))  %>%
  mutate( # This parses the coordinate out of GRID_ID
    y =   0.01 * as.numeric(substr(GRID_ID, 1, 5)),
    x = - 0.01 * as.numeric(substr(GRID_ID, 6, 10)))

# Turn it into a raster
nonhurr_ny <- rasterFromXYZ(sf_ny[, c("x", "y", "NONHURR_FCTR")], crs = crs('+proj=longlat +datum=WGS84'))

# Set up a color palette
pal <- colorBin(
  palette = c("#2c7bb6", "#abd9e9","#ffffbf","#fdae61","#d7191c"),
  domain = values(nonhurr_ny),
  bins = c(0.75, 1.0, 1.25, 1.5, 1.75),
  na.color = "transparent"
)

# Useful for seeing how frequent factors are
# Override digits since it rounds to integers by default
# freq(nonhurr_ny, digits = 3)
leaflet() %>%
  addTiles() %>%
  addRasterImage(nonhurr_ny, colors = pal, opacity = 0.5) %>%
  addLegend(pal = pal, values = values(nonhurr_ny), title = "Non-Hurricane Factor")
```

## State Farm Fire And Casualty Company (25143) - Hurricane Factors {background-color="#A1D6F4"}

```{r State Farm NY hurricane factors}
#| echo: true
#| code-fold: true
#| code-summary: "Show code"


# Turn it into a raster
hurr_ny <- rasterFromXYZ(sf_ny[, c("x", "y", "HURR_FCTR")], crs = crs('+proj=longlat +datum=WGS84'))

# Set up a color palette
pal <- colorBin(
  palette = c("#2c7bb6", "#abd9e9","#ffffbf","#fdae61","#d7191c"),
  domain = values(hurr_ny),
  bins = c(0, 0.5, 1.5, 2.0, 2.5, 3.0),
  na.color = "transparent"
)

leaflet() %>%
  addTiles() %>%
  addRasterImage(hurr_ny, colors = pal, opacity = 0.5) %>%
  addLegend(pal = pal, values = values(hurr_ny), title = "Hurricane Factor")
```
